---
title: "Comparing Contractile Markers and Creating Paper Image"
author: "Arjun Lakshmanan"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: yes
    theme: journal
    df_print: paged
    code_folding: hide
    highlight: pygments
  pdf_document:
    toc: yes
    toc_depth: '3'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Install Packages 

```{r load new packages, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

knitr::opts_chunk$set(echo = TRUE)
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")}
if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("BiocManager")) {install.packages("BiocManager"); require("BiocManager")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("here")) {install.packages("here"); require("here")}
if (!require("devtools")) {install.packages("devtools"); require("devtools")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")}
if (!require("sjmisc")) {install.packages("sjmisc"); require("sjmisc")}
if (!require("gridExtra")) {install.packages("gridExtra"); require("gridExtra")}
if (!require("gplots")) {install.packages("gplots"); require("gplots")}
if (!require("ggvenn")) {install.packages("ggvenn"); require("ggvenn")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}


set.seed(12345)
#here::set_here()
here()
```



# Objective 

The goal of this file is to compare the transcriptional profile of our Contractile cell dataset (INTACT) with Humphrey's (DKD) as well as creating figures accordingly. 

# Load INTACT Dataset (Collected)

Th

```{r echo=TRUE, warning=FALSE, error=FALSE, results = TRUE, message=FALSE, fig.align = 'center'}

population_colors <- c(
  "Pericyte" = "#66C2A5",
  "Efferent VSMC" = "#FC8D62",
  "Afferent VSMC" = "#8DA0CB",
  "Renin Cell" = "#E78AC3",
  "EG Mesangial Cell" = "#A6D854",
  "IG Mesangial Cell" = "#FFD92F"
)

light_population_colors <- c(
  "Pericyte" = "#A8D8C5",       # Lighter Pericyte
  "Efferent VSMC" = "#FDBF96",    # Lighter Efferent VSMC
  "Afferent VSMC" = "#B8C9E1",    # Lighter Afferent VSMC
  "Renin Cell" = "#F2A4D6",       # Lighter Renin Cell
  "EG Mesangial Cell" = "#B9E09B", # Lighter EG Mesangial Cell
  "IG Mesangial Cell" = "#FFEB88"  # Lighter IG Mesangial Cell
)

very_light_population_colors <- c(
  "Pericyte" = "#D4EDE4",       # Very light Pericyte
  "Efferent VSMC" = "#FFE0CC",  # Very light Efferent VSMC
  "Afferent VSMC" = "#DEE5F2",  # Very light Afferent VSMC
  "Renin Cell" = "#F8D3E8",     # Very light Renin Cell
  "EG Mesangial Cell" = "#DDF2C7", # Very light EG Mesangial Cell
  "IG Mesangial Cell" = "#FFF4C2"  # Very light IG Mesangial Cell
)

colorblind_friendly_population_colors <- c(
  "Pericyte" = "#56B4E9",        # Blue
  "Efferent VSMC" = "#E69F00",   # Orange
  "Afferent VSMC" = "#009E73",   # Green
  "Renin Cell" = "#F0E442",      # Yellow
  "EG Mesangial Cell" = "#0072B2", # Deep Blue
  "IG Mesangial Cell" = "#D55E00", # Reddish-Orange
  "TAL A" = "#CC79A7",        # Purple (distinct from others)
  "TAL B" = "#882255",       # Dark Red (strong contrast)
  "MD" = "#999999" # Gray (neutral but visible)
)


```


```{r echo=TRUE, warning=FALSE, error=FALSE, results = TRUE, message=FALSE, fig.align = 'center'}


SO.INTACT <- readRDS(here("Datasets", "renamed", "INTACT_JGA_renamed.rds"))
DimPlot(SO.INTACT, group.by = "class.JGA", cols = colorblind_friendly_population_colors)


```


# Load DKD Dataset 

```{r echo=TRUE, warning=FALSE, error=FALSE, results = TRUE, message=FALSE, fig.align = 'center'}

SO.DKD <- readRDS(here("Datasets", "renamed", "DKD_JGA_renamed.rds"))


Idents(SO.DKD) <- SO.DKD@meta.data$class.JGA

DimPlot(SO.DKD, group.by = "class.JGA", cols = light_population_colors, reduction = "umap")

```



# ForLoop INTACT DEG Lists

```{r echo=TRUE, warning=FALSE, error=FALSE, results = TRUE, message=FALSE, fig.align = 'center'}


markers_result <- list()

clusters <- unique(SO.INTACT$class.JGA)

for (i in clusters) {
  
  # Replace spaces with underscores in cluster names
  clean_name <- gsub(" ", "_", i)
  
  # Find markers for the cluster
  markers <- FindMarkers(SO.INTACT, ident.1 = i, group.by = "class.JGA", min.pct = 0.1, logfc.threshold = 0.25)
  markers <- markers %>%
    rownames_to_column(var = "gene") %>%
    arrange(desc(avg_log2FC))
  
  markers_result[[clean_name]] <- markers
  
  # Create variable name without spaces
  var_name <- paste0(clean_name, ".INTACT.markers") 
  assign(var_name, markers, envir = .GlobalEnv)

}



```


# ForLoop DKD DEG Lists

```{r echo=TRUE, warning=FALSE, error=FALSE, results = TRUE, message=FALSE, fig.align = 'center'}

markers_result <- list()

clusters <- unique(SO.DKD$class.JGA)


for (i in clusters) {
  
  clean_name <- gsub(" ", "_", i)
  
  markers <- FindMarkers(SO.DKD, ident.1 = i, group.by = "class.JGA", min.pct = 0.1, logfc.threshold = 0.25)
  markers <- markers %>%
    rownames_to_column(var = "gene") %>%
    arrange(desc(avg_log2FC))
  
  markers_result[[clean_name]] <- markers
  
  var_name <- paste0(clean_name, ".DKD.markers") 
  assign(var_name, markers, envir = .GlobalEnv)

}



```





# Name Lists 


```{r echo=TRUE, warning=FALSE, error=FALSE, results = TRUE, message=FALSE, fig.align = 'center'}

INTACT_list <- list(`Pericyte.INTACT.markers`, `Renin_Cell.INTACT.markers`, `Efferent_VSMC.INTACT.markers`, `Afferent_VSMC.INTACT.markers`, `IG_Mesangial_Cell.INTACT.markers`, `EG_Mesangial_Cell.INTACT.markers`)


DKD_list <- list(`Pericyte.DKD.markers`, `Renin_Cell.DKD.markers`, `Efferent_VSMC.DKD.markers`, `Afferent_VSMC.DKD.markers`, `IG_Mesangial_Cell.DKD.markers`, `EG_Mesangial_Cell.DKD.markers`)


names(INTACT_list) <- c("Pericyte.INTACT.markers", "Renin Cell.INTACT.markers", "Efferent VSMC.INTACT.markers", "Afferent VSMC.INTACT.markers", "IG_Mesangial_Cell.INTACT.markers", "EG_Mesangial_Cell.INTACT.markers")


names(DKD_list) <- c("Pericyte.DKD.markers", "Renin Cell.DKD.markers", "Efferent VSMC.DKD.markers", "Afferent VSMC.DKD.markers", "IG_Mesangial_Cell.DKD.markers", "EG_Mesangial_Cell.DKD.markers")






```




# Scatterplot Forloop + R-squared Table


```{r echo=TRUE, warning=FALSE, error=FALSE, results = TRUE, message=FALSE, fig.align = 'center'}

library(dplyr)
library(ggplot2)
library(ggrepel)
library(broom)

# Initialize table for storing R-squared values and slopes
rsq_table <- data.frame(matrix(ncol = 7, nrow = (length(INTACT_list) * length(DKD_list))))
names(rsq_table) <- c("INTACT", "DKD", "R-squared", "Slope", "Shared_DEGs", "INTACT_UniqueDEGs", "DKD_UniqueDEGs")

# List to store scatterplots
scatterplots <- list()

## START FOR {LOOP} LEVEL 1 ##
for (i in 1:length(INTACT_list)) {
  
  x.markers <- INTACT_list[[i]]
  x.markers_tb <- x.markers %>%
    data.frame() %>%
    as_tibble()
  
  ## START FOR {LOOP} LEVEL 2 ##
  for (j in 1:length(DKD_list)) {
    y.markers <- DKD_list[[j]]
    y.markers_tb <- y.markers %>%
      data.frame() %>%
      as_tibble()
    
    # X-Y DEGs Intersection Table
    xy.comp <- inner_join(x.markers_tb, y.markers_tb, by = "gene")

    # Ensure enough genes exist for regression
    if(nrow(xy.comp) > 1) { 
      
      # Selecting only relevant columns
      xy.comp.R2 <- xy.comp %>%
        dplyr::select(avg_log2FC.x, avg_log2FC.y)
      
      # Correct regression model
      model <- lm(avg_log2FC.y ~ avg_log2FC.x, data = xy.comp.R2)
      
      # Extract slope and correct R² calculation
      slope <- coef(model)["avg_log2FC.x"]
      n_rsq <- summary(model)$r.squared

    } else {
      slope <- NA
      n_rsq <- NA  # Not enough data for a valid regression
    }

    # Calculate DEGs counts
    Shared_DEGs <- nrow(xy.comp)
    INTACT_UniqueDEGs <- nrow(x.markers_tb) - Shared_DEGs
    DKD_UniqueDEGs <- nrow(y.markers_tb) - Shared_DEGs

    # Add to R-squared table
    index <- ((i-1) * length(DKD_list)) + j
    rsq_table[index, 1] = names(INTACT_list)[i]
    rsq_table[index, 2] = names(DKD_list)[j]
    rsq_table[index, 3] = n_rsq
    rsq_table[index, 4] = slope  # Adding the slope to the table
    rsq_table[index, 5] = Shared_DEGs
    rsq_table[index, 6] = INTACT_UniqueDEGs
    rsq_table[index, 7] = DKD_UniqueDEGs

    # Generate Scatterplot for each pair
    p <- ggplot(xy.comp, aes(x = avg_log2FC.x, y = avg_log2FC.y, label = gene)) +
      geom_point(color = "#636EFA", alpha = 0.6, size = 2.5) +  # Blue dots with transparency
      geom_smooth(method = "lm", color = "#1E90FF", linetype = "dashed", se = FALSE) + # Best-fit regression line
      geom_text_repel(size = 4, segment.size = 0.2, segment.color = "grey50") +  # Gene labels
      annotate("text", x = min(xy.comp$avg_log2FC.x) + 0.2, 
               y = max(xy.comp$avg_log2FC.y) - 0.2, 
               label = paste0("R² = ", round(n_rsq, 3), "\nSlope = ", round(slope, 3)), 
               color = "black", size = 6, hjust = 0) +
      labs(
        title = paste(names(INTACT_list)[i], "vs", names(DKD_list)[j]),
        x = paste("avg_log2FC -", names(INTACT_list)[i]),
        y = paste("avg_log2FC -", names(DKD_list)[j])
      ) +
      theme_minimal(base_size = 14) +
      theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title = element_text(face = "bold"),
        axis.line = element_line(color = "black", size = 1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
      )

    # Store the plot
    scatterplots[[paste(names(INTACT_list)[i], names(DKD_list)[j], sep = "_vs_")]] <- p
  }
}

# Sort and print the R-squared table
rsq_table <- rsq_table %>%
  arrange(desc(`R-squared`))

print(rsq_table)

# Display all scatterplots
for (plot_name in names(scatterplots)) {
  print(scatterplots[[plot_name]])
}



```




# Create HeatMap of Datasets 


```{r echo=TRUE, warning=FALSE, error=FALSE, results = TRUE, message=FALSE, fig.align = 'center'}



df <- rsq_table %>%
                dplyr::select(INTACT, DKD, "R-squared") %>%
                mutate(INTACT = sub("_Named_", " ", INTACT)) %>%
                mutate(DKD = sub("_Named_", " ", DKD)) %>%
                pivot_wider(names_from = DKD, values_from = "R-squared") %>%
                column_to_rownames(var = "INTACT") %>%
                as.matrix()


library(corrplot)
## Correlation Plot
corrplot(df, method = 'color')




DimPlot(SO.INTACT, group.by = "class.JGA")




```




# Specific Marker Expression 







# Figures 

## First Panel 

```{r echo=TRUE, warning=FALSE, error=FALSE, results = TRUE, message=FALSE, fig.align = 'center'}


p1 <- DimPlot(SO.DKD, reduction = "umap", cols = very_light_population_colors)  + 
  ylab("UMAP 2") +
  xlab("UMAP 1") +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 20),
    axis.line = element_line(size = 1, colour = "black"),
    axis.text = element_blank(),          # Remove axis text
    axis.ticks = element_blank(),         # Optional: Remove axis ticks
    text = element_text(size = 20)
  ) + ggtitle("UMAP - DKD Contractile Clusters") 


p2 <- DimPlot(SO.INTACT, reduction = "umap", group.by = "class.JGA", cols = population_colors) + 
  ylab("UMAP 2") +
  xlab("UMAP 1") +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 20),
    axis.line = element_line(size = 1, colour = "black"),
    axis.text = element_blank(),          # Remove axis text
    axis.ticks = element_blank(),         # Optional: Remove axis ticks
    text = element_text(size = 20)
  ) + ggtitle("UMAP - INTACT Contractile Clusters") 

print(p1)
print(p2)




```


## Panel 2


```{r echo=TRUE, warning=FALSE, error=FALSE, results = TRUE, message=FALSE, fig.align = 'center', fig.length = 8, fig.width = 8}

library(patchwork)

library(ggplot2)
library(patchwork)

Idents(SO.DKD) <- SO.DKD@meta.data$class.JGA
Idents(SO.INTACT) <- SO.INTACT@meta.data$class.JGA

Idents(SO.DKD) <- SO.DKD@meta.data$class.JGA
Idents(SO.INTACT) <- SO.INTACT@meta.data$class.JGA


markers.to.plot <- c("Nkain2", 
                     "Col23a1", # Adora1 didn't look good 
                     "Ren1", 
                     "Hpse2", 
                     "Tenm2", 
                     "Tshz2")



#cluster_order <- c("Pericyte", "Efferent VSMC", "Afferent VSMC", "Renin Cell", "EG Mesangial Cell", "IG Mesangial Cell")

#SO.DKD@meta.data$class.JGA <- factor(SO.DKD@meta.data$class.JGA, levels = cluster_order)


p3 <- DotPlot(SO.DKD, features = markers.to.plot) +
  coord_flip() +
  ggtitle("Gene Expression Across Clusters - DKD") + 
  xlab("Clusters") + ylab("Marker Genes") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size = 14, face = "bold", hjust = 0.5))

p4 <- DotPlot(SO.INTACT, features = markers.to.plot) +
  coord_flip() +
  ggtitle("Gene Expression Across Clusters - INTACT") + 
  xlab("Clusters") + ylab("Marker Genes") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size = 14, face = "bold", hjust = 0.5))

p3
p4



```



## Panel 3

```{r echo=TRUE, warning=FALSE, error=FALSE, results = TRUE, message=FALSE, fig.align = 'center', fig.length = 15, fig.width = 10}


xy.comp <- inner_join(IG_Mesangial_Cell.INTACT.markers, IG_Mesangial_Cell.DKD.markers, by = "gene")

df.upper <- subset(xy.comp, avg_log2FC.x > 1 & avg_log2FC.y > 1)
df.lower <- subset(xy.comp, avg_log2FC.x < -1 & avg_log2FC.y < -1)

model <- lm(avg_log2FC.y ~ avg_log2FC.x, data = xy.comp)

slope <- coef(model)["avg_log2FC.x"]
r_squared <- summary(model)$r.squared



library(ggplot2)
library(ggrepel)

xy.comp <- inner_join(IG_Mesangial_Cell.INTACT.markers, IG_Mesangial_Cell.DKD.markers, by = "gene")

p5 <- ggplot(xy.comp, aes(x = avg_log2FC.x, y = avg_log2FC.y, label = gene)) +
  
  geom_point(color = "#636EFA", alpha = 0.6, size = 2.5) + 
  
  geom_point(data = df.upper, color = "#EF553B", size = 3.5) + 
  geom_point(data = df.lower, color = "#00CC96", size = 3.5) + 
  geom_smooth(method = "lm", color = "#1E90FF", linetype = "dashed", se = FALSE) +
  
  geom_text_repel(data = rbind(df.upper, df.lower), segment.size = 0.2, segment.color = "grey50", 
                  size = 5) +
    annotate("text", x = min(xy.comp$avg_log2FC.x) + 0.2, 
           y = max(xy.comp$avg_log2FC.y) - 0.2, 
           label = paste0("R² = ", round(r_squared, 3), "\nSlope = ", round(slope, 3)), 
           color = "black", size = 6, hjust = 0) +

  
  labs(
    title = "Intraglomerular Mesangial Gene Expression: INTACT vs DKD",
    x = "Average log2FC - INTACT IG Mesangial",
    y = "Average log2FC - DKD IG Mesangial"
  ) +
  
  theme_minimal(base_size = 16) + 
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold"), 
    axis.line = element_line(color = "black", size = 1),  
    panel.grid.major = element_blank(),  
    panel.grid.minor = element_blank()  
  )


p5




```



## Panel 4 


```{r echo=TRUE, warning=FALSE, error=FALSE, results = TRUE, message=FALSE, fig.align = 'center', fig.length = 13, fig.width - 15}


library(dplyr)
library(tidyr)
library(ggplot2)
library(reshape2)
library(RColorBrewer)

rsq_table <- rsq_table %>%
  mutate(
    INTACT = gsub("\\.", " ", INTACT),
    DKD = gsub("\\.", " ", DKD),
    INTACT = gsub("INTACT markers", "", INTACT), 
    DKD = gsub("DKD markers", "", DKD),
    INTACT = gsub("_", " ", INTACT),
    DKD = gsub("_", " ", DKD)
  )

df <- rsq_table %>%
  dplyr::select(INTACT, DKD, `R-squared`) %>%
  pivot_wider(names_from = DKD, values_from = `R-squared`) %>%
  column_to_rownames(var = "INTACT") %>%
  as.matrix()

cor_long <- melt(df)

p6 <- ggplot(cor_long, aes(x = Var1, y = Var2, fill = value)) +
  
  geom_tile(color = "white", size = 0.3) +
  
  geom_text(aes(label = sprintf("%.2f", value)), size = 5, fontface = "bold",
            color = ifelse(cor_long$value > 0.5, "white", "black")) + 
  
  scale_fill_gradientn(colors = rev(brewer.pal(9, "RdBu")), 
                       limits = c(0, 1), name = "R²") +
  
  theme_minimal(base_size = 16) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 14, face = "bold", color = "black"),
    axis.text.y = element_text(size = 14, face = "bold", color = "black"),
    axis.title = element_text(size = 16, face = "bold"),
    panel.grid = element_blank(),  # No distracting grid lines
    legend.position = "right",
    legend.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5)
  ) +
  
  labs(title = "Correlation of Contractile Cell Clusters (INTACT vs DKD)",
       x = "DKD Clusters",
       y = "INTACT Clusters")

p6

## JWN Coding

# Define the desired order
cor_long$Var1 <- trimws(cor_long$Var1)
cor_long$Var2 <- trimws(cor_long$Var2)

# Convert Var1 and Var2 into ordered factors
cor_long$Var1 <- factor(cor_long$Var1, levels = c("Pericyte", "Efferent VSMC", "Afferent VSMC", 
                                                  "Renin Cell", "EG Mesangial Cell", "IG Mesangial Cell"))
cor_long$Var2 <- factor(cor_long$Var2, levels = c("Pericyte", "Efferent VSMC", "Afferent VSMC", 
                                                  "Renin Cell", "EG Mesangial Cell", "IG Mesangial Cell"))

# Plot
p6 <- ggplot(cor_long, aes(x = Var2, y = Var1, fill = value)) + 
  geom_tile(color = "white", size = 0.3) + 
  geom_text(aes(label = sprintf("%.2f", value)), size = 5, fontface = "bold",
            color = ifelse(cor_long$value > 0.5, "white", "black")) + 
  scale_fill_gradientn(colors = rev(brewer.pal(9, "RdBu")), 
                       limits = c(0, 1), name = "R²") +
  theme_minimal(base_size = 16) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 14, face = "bold", color = "black"),
    axis.text.y = element_text(size = 14, face = "bold", color = "black"),
    axis.title = element_text(size = 16, face = "bold"),
    panel.grid = element_blank(),
    legend.position = "right",
    legend.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5)
  ) +
  labs(title = "Correlation between INTACT vs DKD",
       x = "DKD Clusters",
       y = "INTACT Clusters") +
  coord_fixed(ratio = .5)

p6


```


# Final Figure

```{r echo=TRUE, warning=FALSE, error=FALSE, results = TRUE, message=FALSE, fig.align = 'center'}


p1
p2
p3
p4
p5
p6

```




# Session Info

```{r sessioninfo, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

sessionInfo()

```


